---
title: "10  Visual Statistical Analysis"
author: "SU, Bo-Han"
---

# **10.1 Learning Outcome**

------------------------------------------------------------------------

In this hands-on exercise, you will gain hands-on experience on using:

-   ggstatsplot package to create visual graphics with rich statistical information,

-   performance package to visualise model diagnostics, and

-   parameters package to visualise model parameters

# **10.2 Visual Statistical Analysis with ggstatsplot**

------------------------------------------------------------------------

[**ggstatsplot**](https://indrajeetpatil.github.io/ggstatsplot/index.html) is an extension of [**ggplot2**](https://ggplot2.tidyverse.org/) package for creating graphics with details from statistical tests included in the information-rich plots themselves.

```         
-   To provide alternative statistical inference methods by default.
-   To follow best practices for statistical reporting. For all statistica
```

# **10.3 Getting Started**

------------------------------------------------------------------------

## **10.3.1 Installing and launching R packages**

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## **10.3.2 Importing data**

```{r}
exam <- read_csv("data/Exam_data.csv")
```

```{r}
exam
```

## **10.3.3 One-sample test: *gghistostats()* method**

In the code chunk below, [*gghistostats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/gghistostats.html) is used to to build an visual of one-sample test on English scores.

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## **10.3.5 How to interpret Bayes Factor**

------------------------------------------------------------------------

A **Bayes Factor** can be any positive number. One of the most common interpretations is this one—first proposed by Harold Jeffereys (1961) and slightly modified by [Lee and Wagenmakers](https://www-tandfonline-com.libproxy.smu.edu.sg/doi/pdf/10.1080/00031305.1999.10474443?needAccess=true) in 2013:

## **10.3.6 Two-sample mean test: *ggbetweenstats()***

## ---

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for two-sample mean test of Maths scores by gender.

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

## **10.3.7 Oneway ANOVA Test: *ggbetweenstats()* method**

------------------------------------------------------------------------

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for One-way ANOVA test on English score by race.

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

-   “ns” → only non-significant

-   “s” → only significant

-   “all” → everything

## **10.3.8 Significant Test of Correlation: *ggscatterstats()***

## ---

In the code chunk below, [*ggscatterstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggscatterstats.html) is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## **10.3.9 Significant Test of Association (Depedence) : *ggbarstats()* methods**

## ---

In the code chunk below, the Maths scores is binned into a 4-class variable by using [*cut()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut).

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

In this code chunk below [*ggbarstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbarstats.html) is used to build a visual for Significant Test of Association

```{r}
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```
