---
title: "2 Beyond ggplot2 Fundamentals"
author: "SU, Bo-Han"
---

# **2.1 Overview**

------------------------------------------------------------------------

In this chapter, you will be introduced to several ggplot2 extensions for creating more elegant and effective statistical graphics. By the end of this exercise, you will be able to:

-   control the placement of annotation on a graph by using functions provided in ggrepel package,

-   create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages,

-   plot composite figure by combining ggplot2 graphs by using patchwork package.

# **2.2 Getting started**

------------------------------------------------------------------------

## 2.2.1 Installing and loading the required libraries

In this exercise, beside tidyverse, four R packages will be used. They are:

-   ggrepel: an R package provides geoms for ggplot2 to repel overlapping text labels.

-   ggthemes: an R package provides some extra themes, geoms, and scales for ‘ggplot2’.

-   hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.

-   patchwork: an R package for preparing composite figure created using ggplot2.

Code chunk below will be used to check if these packages have been installed and also will load them onto your working R environment.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

### **2.2.2 Importing data**

For the purpose of this exercise, a data file called *Exam_data* will be used. It consists of year end examination grades of a cohort of primary 3 students from a local school. It is in csv file format.

The code chunk below imports *exam_data.csv* into R environment by using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html)function of [**readr**](https://readr.tidyverse.org/) package. **readr** is one of the tidyverse package.

```{r}
library(tidyverse)
exam_data <- read_csv("data/Exam_data.csv")
```

There are a total of seven attributes in the exam_data tibble data frame. Four of them are categorical data type and the other three are in continuous data type.

-   The categorical attributes are: ID, CLASS, GENDER and RACE.

-   The continuous attributes are: MATHS, ENGLISH and SCIENCE.

# **2.3 Beyond ggplot2 Annotation: ggrepel**

------------------------------------------------------------------------

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.3.1 Working with ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

# 2.4 Beyond ggplot2 Themes

------------------------------------------------------------------------

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

## 2.4.1 Working with ggtheme package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## 2.4.2 Working with hrbthems package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

# 2.5 Beyond Single Graph

------------------------------------------------------------------------

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
p1
```

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
p2
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
p3
```

## 2.5.1 Creating Composite Graphics: pathwork methods

## 2.5.2 Combining two ggplot2 graphs

```{r}
p1+p2
```

## 2.5.3 Combining three ggplot2 graphs

```{r}
p1+p2+p3
```

## 2.5.4 Creating a composite figure with tag

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

## 2.5.5 Crating figure with insert

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

## 2.5.6 Creating a composite figure by using pathwork and ggtheme

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

# **2.6 Reference**

------------------------------------------------------------------------

-   [Patchwork R package goes nerd viral](https://www.littlemissdata.com/blog/patchwork)

-   [ggrepel](https://ggrepel.slowkow.com/)

-   [ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

-   [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)

-   [ggplot tips: Arranging plots](https://albert-rapp.de/post/2021-10-28-extend-plot-variety/)

-   [ggplot2 Theme Elements Demonstration](https://henrywang.nl/ggplot2-theme-elements-demonstration/)

-   [ggplot2 Theme Elements Reference Sheet](https://isabella-b.com/blog/ggplot2-theme-elements-reference/)
